Measurement and Probability in Fuzzy Quantum Theories

### 2017-02-02 prop.tex
prop.tex \section{Quantum and Classical} Modifying words...

\begin{lemma}Given an orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$
for a given Hilbert space~$\Hilb$ of dimension~$d$, and a quantum
IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$. Then, the function
$\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$ defined by
$\bar{\mu}^{\Omega}\left(E\right)=\bar{\mu}\left(\sum_{\ket{j}\in E}\proj{j}\right)$
is a classical IVPM, where $\sum_{\ket{j}\in\emptyset}\proj{j}$ is
defined to be $\mathbb{0}$.\end{lemma}
\begin{proof}
~
\begin{itemize}
\item $\bar{\mu}^{\Omega}(\emptyset)=\bar{\mu}\left(\sum_{\ket{j}\in\emptyset}\proj{j}\right)=\bar{\mu}\left(\mathbb{0}\right)=[0,0]$. 
\item $\bar{\mu}^{\Omega}(\Omega)=\bar{\mu}\left(\sum_{\ket{j}\in\Omega}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}\right)=[1,1]$. 
\item For any event $E$, $\bar{\mu}^{\Omega}\left(\Omega\backslash E\right)=\bar{\mu}\left(\sum_{\ket{j}\in\Omega\backslash E}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}-\sum_{\ket{j}\in E}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}\right)-\bar{\mu}\left(\sum_{\ket{j}\in E}\proj{j}\right)=\left[1,1\right]-\bar{\mu}^{\Omega}\left(E\right)$. 
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, we have $\bar{\mu}^{\Omega}\left(\bigcup_{i=1}^{N}E_{i}\right)=\bar{\mu}\left(\sum_{\ket{j}\in\bigcup_{i=1}^{N}E_{i}}\proj{j}\right)=\bar{\mu}\left(\sum_{i=1}^{N}\sum_{\ket{j}\in E_{i}}\proj{j}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(\sum_{\ket{j}\in E_{i}}\proj{j}\right)=\sum_{i=1}^{N}\bar{\mu}\left(E_{i}\right)$. 
\end{itemize}
\end{proof}
The previous lemma suggests that given an orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$,
if we only consider the events generated by $\Omega$, $\events^{\Omega}=\set{\sum_{\ket{j}\in E}\proj{j}}{E\subseteq\Omega}$,
then any quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
behaves classically when restricting the domain on $\events'$, i.e.,
consider $\bar{\mu}|_{\events^{\Omega}}$. In particular, if $\bar{\mu}$
is convex, each $\bar{\mu}|_{\events^{\Omega}}$ has a non-empty core.
However, when they are put together, there is no quantum probability
measure~$\mu:\events\rightarrow\left[0,1\right]$ whose restriction
$\mu|_{\events^{\Omega}}$ inside all the core of $\bar{\mu}|_{\events^{\Omega}}$.
Notice that every events in $\events^{\Omega}$ are commuting, which
consistent with the idea of the Kochen-Specker theorem, that each
triple of commuting observables could be colored well, but we cannot
color all observables which are not all commuting.\footnote{\yutsung{Since Bell's theorem only applies to entangled states, the
exists of entangled states is another reason why quantum theory is
different from a classical theory. However, defining product quantum
IVPM is much harder, and beyond the discussion here.}}

\begin{lemma}Given a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$,
if $\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$ is a classical
convex IVPM for any orthonormal basis~$\Omega$, then $\bar{\mu}$
is a quantum convex IVPM.\end{lemma}
\begin{proof}
Pick two commuting events~$P_{0},P_{1}\in\events$. Since they are
commuting, they can be diagonalized by a common orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$.
Since they are projectors, the eigenvalues of them are $0$ or $1$
so that we can express them as $P_{0}=\sum_{\ket{j}\in E_{0}}\proj{j}$
and $P_{1}=\sum_{\ket{j}\in E_{1}}\proj{j}$. Then, 
\begin{eqnarray*}
 &  & \bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\\
 & = & \bar{\mu}\left(\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)\vee\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\right)+\bar{\mu}\left(\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\right)\\
 & = & \bar{\mu}\left(\sum_{\ket{j}\in E_{0}\cup E_{1}}\proj{j}\right)+\bar{\mu}\left(\sum_{\ket{j}\in E_{0}\cap E_{1}}\proj{j}\right)\\
 & = & \bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\\
 & \subseteq & \bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\\
 & = & \bar{\mu}\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)+\bar{\mu}\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\\
 & = & \bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)
\end{eqnarray*}
\end{proof}
\begin{lemma}When the number of elements in $\Omega$ is less than
or equal to 3, every classical convex IVPM~$\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$.\end{lemma}
\begin{proof}
Given two events~$E_{0},E_{1}\subseteq\Omega$, we want to verify
\begin{equation}
\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\textrm{ .}\label{eq:classical-convex}
\end{equation}
When $E_{0}\cap E_{1}=\emptyset$, equation~(\ref{eq:classical-convex})
is a special case of equation~(\ref{eq:classical-include}). When
$E_{0}\cup E_{1}=\Omega$, we have $E_{0}^{c}\cap E_{1}^{c}=\emptyset$,
where $E_{0}^{c}=\Omega\backslash E_{0}$ and $E_{1}^{c}=\Omega\backslash E_{1}$.
Thus, 
\begin{eqnarray*}
 &  & \bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\
 & \Rightarrow & \left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\
 & \Rightarrow & \bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cup E_{1}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cap E_{1}^{c}\right)\right)\subseteq\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{1}^{c}\right)\right)\\
 & \Rightarrow & \bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\textrm{ .}
\end{eqnarray*}
When the number of elements in $\Omega$ is less than or equal to
3, either $E_{0}\cap E_{1}=\emptyset$ or $E_{0}\cup E_{1}=\Omega$
must happen so that $\bar{\mu}^{\Omega}$is convex.
\end{proof}
\begin{thm}Given a Hilbert space~$\Hilb$ of dimension~$3$, every
quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
is convex.\end{thm}
\begin{proof}
The direct consequence of the previous lemmas.
\end{proof}

>>>>

\begin{lemma}Given an orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$
for a given Hilbert space~$\Hilb$ of dimension~$d$, and a quantum
IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$. Then, the function
$\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$ defined by
$\bar{\mu}^{\Omega}\left(E\right)=\bar{\mu}\left(\sum_{\ket{j}\in E}\proj{j}\right)$
is a classical IVPM, where $\sum_{\ket{j}\in\emptyset}\proj{j}$ is
defined to be $\mathbb{0}$.\end{lemma}
\begin{proof}
~
\begin{itemize}
\item $\bar{\mu}^{\Omega}(\emptyset)=\bar{\mu}\left(\sum_{\ket{j}\in\emptyset}\proj{j}\right)=\bar{\mu}\left(\mathbb{0}\right)=[0,0]$. 
\item $\bar{\mu}^{\Omega}(\Omega)=\bar{\mu}\left(\sum_{\ket{j}\in\Omega}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}\right)=[1,1]$. 
\item For any event $E$, $\bar{\mu}^{\Omega}\left(\Omega\backslash E\right)=\bar{\mu}\left(\sum_{\ket{j}\in\Omega\backslash E}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}-\sum_{\ket{j}\in E}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}\right)-\bar{\mu}\left(\sum_{\ket{j}\in E}\proj{j}\right)=\left[1,1\right]-\bar{\mu}^{\Omega}\left(E\right)$. 
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, we have $\bar{\mu}^{\Omega}\left(\bigcup_{i=1}^{N}E_{i}\right)=\bar{\mu}\left(\sum_{\ket{j}\in\bigcup_{i=1}^{N}E_{i}}\proj{j}\right)=\bar{\mu}\left(\sum_{i=1}^{N}\sum_{\ket{j}\in E_{i}}\proj{j}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(\sum_{\ket{j}\in E_{i}}\proj{j}\right)=\sum_{i=1}^{N}\bar{\mu}\left(E_{i}\right)$. 
\end{itemize}
\end{proof}
The previous lemma suggests that given an orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$,
if we only consider the events generated by $\Omega$, $\events^{\Omega}=\set{\sum_{\ket{j}\in E}\proj{j}}{E\subseteq\Omega}$,
then any quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
behaves classically when restricting the domain on $\events^{\Omega}$,
i.e., consider $\bar{\mu}|_{\events^{\Omega}}$. In particular, if
$\bar{\mu}$ is convex, each $\bar{\mu}|_{\events^{\Omega}}$ has
a non-empty core. However, the core elements of each $\bar{\mu}|_{\events^{\Omega}}$cannot
be glued together, and gives a core element of $\bar{\mu}$. The inability
of gluing core elements among commuting $\events^{\Omega}$ to non-commuting
$\events$ is consistent with the idea of the Kochen-Specker theorem,
that each triple of commuting observables could be colored well, but
we cannot color all observables which are not all commuting.\footnote{\yutsung{Since Bell's theorem only applies to entangled states, the
exists of entangled states is another reason why quantum theory is
different from a classical theory. However, defining product quantum
IVPM is much harder, and beyond the discussion here.}}

\begin{lemma}Given a quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$,
if $\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$ are all
convex for any orthonormal basis~$\Omega$, then $\bar{\mu}$ is
convex.\end{lemma}
\begin{proof}
Pick two commuting events~$P_{0},P_{1}\in\events$. Since they are
commuting, they can be diagonalized by a common orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$.
Since they are projectors, the eigenvalues of them are $0$ or $1$
so that we can express them as $P_{0}=\sum_{\ket{j}\in E_{0}}\proj{j}$
and $P_{1}=\sum_{\ket{j}\in E_{1}}\proj{j}$, where $E_{0},E_{1}\subseteq\Omega$.
Then, 
\begin{eqnarray*}
 &  & \bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\\
 & = & \bar{\mu}\left(\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)\vee\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\right)+\bar{\mu}\left(\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\right)\\
 & = & \bar{\mu}\left(\sum_{\ket{j}\in E_{0}\cup E_{1}}\proj{j}\right)+\bar{\mu}\left(\sum_{\ket{j}\in E_{0}\cap E_{1}}\proj{j}\right)\\
 & = & \bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\\
 & \subseteq & \bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\\
 & = & \bar{\mu}\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)+\bar{\mu}\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\\
 & = & \bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)
\end{eqnarray*}
\end{proof}
\begin{lemma}When the number of elements in $\Omega$ is less than
or equal to 3, every classical IVPM~$\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$
is convex.\end{lemma}
\begin{proof}
Given two events~$E_{0},E_{1}\subseteq\Omega$, we want to verify
\begin{equation}
\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\textrm{ .}\label{eq:classical-convex}
\end{equation}
When $E_{0}\cap E_{1}=\emptyset$, equation~(\ref{eq:classical-convex})
is a special case of equation~(\ref{eq:classical-include}). When
$E_{0}\cup E_{1}=\Omega$, we have $E_{0}^{c}\cap E_{1}^{c}=\emptyset$,
where $E_{0}^{c}=\Omega\backslash E_{0}$ and $E_{1}^{c}=\Omega\backslash E_{1}$.
Thus, 
\begin{eqnarray*}
 &  & \bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\
 & \Rightarrow & \left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\
 & \Rightarrow & \bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cup E_{1}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cap E_{1}^{c}\right)\right)\subseteq\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{1}^{c}\right)\right)\\
 & \Rightarrow & \bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\textrm{ .}
\end{eqnarray*}
When the number of elements in $\Omega$ is less than or equal to
3, either $E_{0}\cap E_{1}=\emptyset$ or $E_{0}\cup E_{1}=\Omega$
must happen so that $\bar{\mu}^{\Omega}$is convex.
\end{proof}
\begin{thm}Given a Hilbert space~$\Hilb$ of dimension~$3$, every
quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ is convex.\end{thm}
\begin{proof}
The direct consequence of the previous lemmas.
\end{proof}
